---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# parlSMCFCS

The [`mice::parlmice`](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html) alternative for the [smcfcs](https://cran.r-project.org/web/packages/smcfcs/index.html) package: a basic wrapper that allows to run the imputations in parallel. 

This can save a substantial amount of computing time, particularly in cases when continuous variables are imputed and rejection sampling is needed.

## Installation

You can install the package with:

``` r
devtools::install_github("edbonneville/parlSMCFCS")
```

## Example

Load the library

```{r load_lib, eval=FALSE}
library(parlSMCFCS)
```

```{r load_lib_devtools, echo=FALSE, message=FALSE, warning=FALSE}
devtools::load_all()
library(ggplot2)

theme_set(theme_bw(base_size = 14))
```

Check number of cores

```{r check_cores}
# Detect number of cores
parallel::detectCores() 
```

Run the imputations on 3 cores, plot

```{r example, eval=FALSE}
imps <- parlSMCFCS::parlsmcfcs(
  seed = 2021,
  n_cores = 3, # number of cores - 1
  originaldata = smcfcs::ex_compet,
  m = 10,
  smtype = "compet",
  smformula = list(
    "Surv(t, d == 1) ~ x1 + x2",
    "Surv(t, d == 2) ~ x1 + x2"
  ),
  numit = 50,
  method = c("", "", "norm", "norm")
)

plot(imps)
```


```{r convergence_plot, echo=FALSE}
imps <- readRDS("benchmarking/readme_imps.rds")
plot(imps)
```


## Benchmarks

(Figure to appear below)

```{r benchmarking, echo=FALSE, eval=FALSE}
# read the rds file, then plot
# benches <- readRDS(...)

ggplot(
  data = benches,
  aes(x = m, y = Time, colour = Cores, linetype = Cores)
) +
  geom_line(size = 1.5) +
  labs(x = "Number of imputations (m)", y = "Time (seconds)")
```

