---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Rmd settings
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "svg",
  fig.retina = 3, 
  fig.align = "center",
  fig.width = 8,
  fig.height = 5
)

# Load packages
devtools::load_all()
library(ggplot2)

# Set ggplot2 theme
theme_set(
  theme_bw(base_size = 14) +
    theme(
      panel.background = element_rect(fill = "transparent"),
      plot.background = element_rect(fill = "transparent", colour = NA), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      legend.background = element_rect(fill = "transparent", colour = NA), 
      legend.box.background = element_rect(fill = "transparent", colour = NA) 
  )
)
```

# parlSMCFCS

The [`mice::parlmice()`](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html) alternative for the [smcfcs](https://cran.r-project.org/web/packages/smcfcs/index.html) package: a basic wrapper that allows to run the substantive model compatible imputations in parallel. 

It uses the parallel package as backend, and supports a progress bar through the use of the [pbapply](https://github.com/psolymos/pbapply) package.

## Installation

You can install the package with:

``` r
devtools::install_github("edbonneville/parlSMCFCS")
```

Otherwise, you can simply download the single [R script](R/parlsmcfcs.R) and source it directly (assuming dependencies have been installed).

## Example

```{r load_lib, eval=FALSE}
library(parlSMCFCS)
library(ggplot2)
```

First, check the number of available cores on your machine:

```{r check_cores}
# Detect number of cores
parallel::detectCores() 
```

Then, you can run the imputations with the exact same arguments as [`smcfcs::smcfcs()`](https://www.rdocumentation.org/packages/smcfcs/versions/1.4.2/topics/smcfcs). The only additional arguments to include are

- `n_cores`: the number of cores to use. It is recommended to leave at least one free core on your machine for other tasks.
- `cl_type`: the type of cluster to start, either "PSOCK" (recommended for windows) or "FORK" (recommended for Linux/Mac).
- `seed`: optional seed value to reproduce the analysis. If run in parallel this is set with `parallel::clusterSetRNGStream(seed)`, otherwise with `n_cores = 1` it is simply set as `set.seed(seed)`.

Additional optional arguments include

- `m_per_core`: specifies number of imputed datasets per core, should you want to override the default of speading `m` across the `n_cores` as equally as possible.
- `outfile`: path to a .txt file to store any printed output, such as rejection sampling warnings.

We run `m = 10` imputations on the [`smcfcs::ex_compet`](https://www.rdocumentation.org/packages/smcfcs/versions/1.4.2/topics/ex_compet) dataset from the smcfcs package, which has two continuous covariates with missing data with a competing risks substantive model. We run them on 3 cores as follows

```{r example, eval=FALSE}
imps <- parlSMCFCS::parlsmcfcs(
  seed = 2021,
  n_cores = 3, # number of cores - 1
  originaldata = smcfcs::ex_compet,
  m = 10,
  smtype = "compet",
  smformula = list(
    "Surv(t, d == 1) ~ x1 + x2",
    "Surv(t, d == 2) ~ x1 + x2"
  ),
  numit = 50,
  method = c("", "", "norm", "norm")
)

plot(imps)
```


```{r convergence_plot, echo=FALSE}
imps <- readRDS("benchmarking/readme_imps.rds")
plot(imps) + scale_color_brewer(palette = "Set3")
```


## Benchmarks

On the same [`smcfcs::ex_compet`](https://www.rdocumentation.org/packages/smcfcs/versions/1.4.2/topics/ex_compet) dataset, benchmarks were run for different `m` and `n_cores` using the [microbenchmark](https://github.com/joshuaulrich/microbenchmark) package with 5 replications. These were run on [SLURM](https://slurm.schedmd.com/documentation.html) cluster with `cl_type = "FORK"`. Below the mean computation times are plotted

```{r benchmarking, echo=FALSE, eval=TRUE}
# Read the rds file, then plot
benches <- readRDS("benchmarking/timings_data.rds")

ggplot(
  data = benches,
  aes(x = m, y = median_time, colour = Cores, linetype = Cores)
) +
  geom_line(size = 1.5) +
  labs(x = "Number of imputations (m)", y = "Time (seconds)") +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "top")
```

